Board Game Recommendation

### 2015-03-10 paper.Rnw
Added some bibliography and some clarifications
Added some bib and some clarification on some sections. Most work has
been touching up the application.

\hh{Give an example of how the data that you have scraped looks like. This would be a good point to introduce a bit of the notation that you are using below, like the $y_{u,i}$. }
\sam{I still need to tie this back to the modeling section.}

@@

There are a few things that make sense to model first. Certain games are more popular than other games, and certain people will naturally be more or less generous than others in their ratings. These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games. We will implement these in the same way that BigChaos implemented the Global Effects model (page 6, BigChaos paper \hh{which of the BigChaos papers are you refering to here, exactly? At some point you need to include the exact reference in the bib tex, but I'd just like to know for myself ...}). While they implemented 14 steps, we will only use the 2 \hh{which 2? you need to include a few more details here}, due to the complexity of the calculations required and the time with which it is practical to fit a model.

@@

To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting. The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$. Either way, the structure is the same.
\begin{equation}
\hat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
\end{equation}

%  The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user. The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
% 
% To begin $r_{u,i}$ will denote the rating by user $u$ for game $i$, where $y_{u,i}$ is a number between 1 and 10. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
% 
% However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
% 
% \begin{center}
% \begin{align}
% y_{u,i} &= \mu_{global} + error_{global} \\
% error_{global} &= \mu_{i} + error_{game} \\
% error_{game} &= \mu_{u} + error_{user} 
% \end{align}
% \end{center}
% 
% And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
% 
% \begin{center}
% \begin{align}
% \mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
% \end{align}
% \end{center}
% \hh{something went wrong in the formula - what are you summing over?}

\hh{This is a linear approach to estimate ratings -- one of the problems is, that rating is on a finite scale, i.e. it is possible that predictions go outside this interval - what do you do with those predictions?}

>>>>

@@

There are a few things that make sense to model first. Certain games are more popular than other games, and certain people will naturally be more or less generous than others in their ratings. These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games. We will implement these in a similar manner to the way that BigChaos implemented their Global Effects model \citep{bigchaos}. While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others. From the RMSE of their probe set, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%. Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.

@@

To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting (motivation for using $\alpha$ in this manner can be found in \cite{scalecollab}). The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$. Either way, the structure is the same.
\begin{equation}
\hat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
\end{equation}

Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closes interval endpoint.
%  The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user. The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
% 
% To begin $r_{u,i}$ will denote the rating by user $u$ for game $i$, where $y_{u,i}$ is a number between 1 and 10. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
% 
% However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
% 
% \begin{center}
% \begin{align}
% y_{u,i} &= \mu_{global} + error_{global} \\
% error_{global} &= \mu_{i} + error_{game} \\
% error_{game} &= \mu_{u} + error_{user} 
% \end{align}
% \end{center}
% 
% And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
% 
% \begin{center}
% \begin{align}
% \mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
% \end{align}
% \end{center}
% \hh{something went wrong in the formula - what are you summing over?}
